\section{Governed Control Plane}\label{governed-control-plane}

The system is a governed harness for autonomous ML experimentation. The manager proposes an experimental change; the worker materialises a patch and run artifacts; the control plane validates scope, executes the lifecycle, parses metrics, decides keep/discard/failed-invalid, updates memory, and appends the authoritative ledger record. Managers and workers are replaceable, but managers cannot commit trial state.

\subsection{Alignment with Agent Evaluation Criteria}\label{alignment-with-agent-evaluation-criteria}

The KDD Agentic AI Evaluation setting asks how to measure agent behaviour in realistic, stochastic task environments rather than only final task success. Table~\ref{tab:kdd-alignment} maps those concerns to the harness mechanisms and metrics used in this paper.

\begin{table*}[t]
\small
\caption{Mapping workshop evaluation concerns to harness mechanisms and paper metrics.}
\label{tab:kdd-alignment}
\begin{tabular}{@{}
  >{\raggedright\arraybackslash}p{0.28\linewidth}
  >{\raggedright\arraybackslash}p{0.38\linewidth}
  >{\raggedright\arraybackslash}p{0.22\linewidth}@{}}
\toprule
Evaluation concern & Harness mechanism & Reported metric or evidence \\
\midrule
Stochastic or repeated poor agent behaviour & Append-only memory plus repeated-bad detector over edit targets and mechanisms & Repeated-bad rate in Section~5.3 \\
No ground truth for intermediate choices & Decision authority separated from the manager; every decision receives provenance IDs & Provenance completeness in Section~5.4 \\
Deployment and scope safety & NodeSpec editable-scope whitelist, no-op guard, and pending-trial guard & Invalid-action rate and stress trials in Section~5.2 \\
Long-horizon failure accumulation & Explicit lifecycle state machine with terminal \texttt{kept}, \texttt{discarded}, and \texttt{failed\_invalid} outcomes & Failure taxonomy and decision breakdown \\
Benchmark reproducibility & NodeSpec YAML, deterministic patch bridge for structured edits, seeded baseline replicates & Artifact manifest and reproducibility discussion \\
\bottomrule
\end{tabular}
\end{table*}

\subsection{Manager/Worker Separation and Replacement Principle}\label{managerworker-separation-and-replacement-principle}

The manager is responsible for proposal generation: a summary, rationale, target file, and intended experimental effect. The worker is responsible for bounded execution: turning that proposal into a patch, running the configured command, and returning artifact references. The control plane is responsible for authority: validity, state transitions, budget accounting, memory updates, provenance IDs, and final decision.

This separation is the replacement principle. A baseline heuristic manager, prompt manager, LangGraph manager, AIDE-style tree-search manager, local worker, Claw-style worker, or future cloud worker can be swapped in without changing the paper-facing governance contract. The append-only ledger remains the source of truth.

\textbf{Worker replication paths.} Two worker implementations cover the full reproducibility spectrum:

\begin{itemize}
\tightlist
\item
  \passthrough{\lstinline!LocalWorker!} (\passthrough{\lstinline!local\_worker.py!}): parses ``Change PARAM from X to Y'' directives from proposal text and applies edits directly with no external dependencies. Suitable for protocol verification without a running LLM or coding agent.
\item
  \passthrough{\lstinline!ClawWorker!} (\passthrough{\lstinline!claw\_worker.py!}): generates a packet JSON and invokes an external coding-agent path for free-form LLM proposals. For structured proposals, \passthrough{\lstinline!ClawWorker!} routes through a deterministic constant-patch path that requires only Ollama, not a coding agent.
\end{itemize}

Both workers implement the same \passthrough{\lstinline!Worker!} protocol and produce ledger records under the same schema; the control plane and paper-facing governance contract are identical regardless of which worker is used.

\subsection{Trial Lifecycle}\label{trial-lifecycle}

Each trial follows the fixed lifecycle shown in Figure~\ref{fig:lifecycle}. The state machine is the unit of evaluation: every opened budget slot must end as a \passthrough{\lstinline!kept!}, \passthrough{\lstinline!discarded!}, or \passthrough{\lstinline!failed\_invalid!} audit object.

\input{sections/fig1_lifecycle}

In pseudocode:

\begin{lstlisting}
for budget_index in campaign_budget:
    proposal = manager.propose(memory_context, node_spec)
    write_pending_guard(campaign_id, trial_id, proposal)
    worker_result = worker.run(proposal, node_spec)
    validity = validate_patch_scope_and_effect(worker_result.patch_ref, node_spec)
    if not validity.ok:
        decision = failed_invalid(validity.failure_category)
    else:
        metrics = parse_metric(worker_result.raw_log_ref)
        decision = keep_if_metric_improves(metrics, current_best)
    append_trial_record(decision, provenance, artifacts)
    update_memory(decision)
    remove_pending_guard()
\end{lstlisting}

\subsection{Scope Enforcement}\label{scope-enforcement}

The ResNet-trigger \passthrough{\lstinline!NodeSpec!} declares \passthrough{\lstinline!train.py!} as the only editable file. Frozen files include \passthrough{\lstinline!prepare.py!}, \passthrough{\lstinline!program.md!}, node dependencies, data files, and artifact directories. The scope validator checks the generated patch before the candidate can be accepted. A patch touching a frozen path is marked \passthrough{\lstinline!failed\_invalid / invalid\_edit\_scope!}, written to the ledger, and prevented from corrupting node state.

\subsection{Pending-Trial Guard}\label{pending-trial-guard}

Before worker execution, the control plane writes a pending guard under \passthrough{\lstinline!experiments/ledgers/!}. The guard makes mid-trial crashes auditable instead of invisible. Recovery tooling can list, inspect, and resolve stale guards by appending a failed-invalid ledger record before deleting the guard. This preserves the invariant that every opened budget slot has a terminal audit object.

\subsection{Append-Only Ledger}\label{append-only-ledger}

The ledger is an append-only JSONL file per campaign. A trial record contains campaign identity, budget index, proposal summary and rationale, patch and raw-log references, parsed metrics, validity status, decision, failure category, provenance IDs, reproducibility hashes, and timing. The ledger is the durable system of record; generated tables and figures are derived artifacts.

\subsection{Memory Modes}\label{memory-modes}

The memory interface has three modes:

\begin{table*}[t]
\small
\begin{tabular}{@{}
  >{\raggedright\arraybackslash}p{(\linewidth - 2\tabcolsep) * \real{0.5000}}
  >{\raggedright\arraybackslash}p{(\linewidth - 2\tabcolsep) * \real{0.5000}}@{}}
\toprule
\begin{minipage}[b]{\linewidth}\raggedright
Mode
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
Manager context
\end{minipage} \\
\midrule
\bottomrule
\passthrough{\lstinline!none!} & No prior-trial context is injected. \\
\passthrough{\lstinline!append\_only\_summary!} & Prior trial summaries, decisions, metrics, and failure categories are injected. \\
\passthrough{\lstinline!append\_only\_summary\_with\_rationale!} & The summary context is augmented with decision rationales, giving the manager targeted failure signals. \\
\end{tabular}
\end{table*}

The ablation asks whether rationale-linked memory reduces repeated poor proposals more than raw history alone.

\subsection{Decision Authority}\label{decision-authority}

We keep the keep/discard decision in the control plane, not delegated to the manager. Rajasekaran et al.~(2026) found that LLM self-evaluation is systematically lenient -- agents confidently praise mediocre work. A deterministic, externally-owned decision criterion addresses this.

The decision rule is intentionally simple: a valid candidate is kept only if the configured metric improves over the current best in the configured direction. A valid but non-improving candidate is \passthrough{\lstinline!discarded / degraded\_metric!}; an invalid candidate is \passthrough{\lstinline!failed\_invalid!} with a specific failure category. The manager may explain a proposal, but it cannot declare it successful.

\subsection{Guides and Sensors}\label{guides-and-sensors}

\textbf{Table S1: Harness guides and sensors for autonomous ML experimentation.}

\begin{table*}[t]
\small
\begin{tabular}{@{}
  >{\raggedright\arraybackslash}p{(\linewidth - 4\tabcolsep) * \real{0.3333}}
  >{\raggedright\arraybackslash}p{(\linewidth - 4\tabcolsep) * \real{0.3333}}
  >{\raggedright\arraybackslash}p{(\linewidth - 4\tabcolsep) * \real{0.3333}}@{}}
\toprule
\begin{minipage}[b]{\linewidth}\raggedright
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
\textbf{Guide (feedforward -- steers before act)}
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
\textbf{Sensor (feedback -- corrects after act)}
\end{minipage} \\
\midrule
\bottomrule
\textbf{Computational} (deterministic, fast) & Node spec defining valid scope; editable-path whitelist; budget cap; scope constraints & Metric parser (\passthrough{\lstinline!val\_bpb!} to \passthrough{\lstinline!val\_auc!}); state machine enforcing transition legality; pending-trial guard detecting crashes \\
\textbf{Inferential} (semantic, LLM-based) & Manager system prompt; memory injection; proposal rationale template & Repeated-bad detector using edit-target and mechanism similarity to flag redundant proposals \\
\end{tabular}
\end{table*}

This table instantiates Böckeler's guides/sensors taxonomy in the ML experimentation setting.

\subsection{Failure Mode Motivation}\label{failure-mode-motivation}

\textbf{Table S2: Agent failure modes and harness controls.}

\begin{table*}[t]
\small
\begin{tabular}{@{}
  >{\raggedright\arraybackslash}p{(\linewidth - 4\tabcolsep) * \real{0.3333}}
  >{\raggedright\arraybackslash}p{(\linewidth - 4\tabcolsep) * \real{0.3333}}
  >{\raggedright\arraybackslash}p{(\linewidth - 4\tabcolsep) * \real{0.3333}}@{}}
\toprule
\begin{minipage}[b]{\linewidth}\raggedright
Agent failure mode
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
ML experimentation analogue
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
Harness fix
\end{minipage} \\
\midrule
\bottomrule
Declares victory too early & Manager stops after one apparently good trial & Budget enforcer plus fixed trial count \\
Leaves broken state & Failed patch leaves repository or node dirty & Pending-trial guard, git before/after checks, reset script \\
Marks feature done prematurely & Manager claims improvement without parseable metric & \passthrough{\lstinline!FAILED\_INVALID!} state plus metric parser \\
Does not know how to run the app & Manager proposes an out-of-scope edit & Node spec plus editable-path whitelist \\
\end{tabular}
\end{table*}

The design treats these failures as expected lifecycle outcomes. Failed trials are audit records, not hidden exceptions.
